### 2023-01-31 doc/research_presentation/research_presentation.tex
presentation

\documentclass[aspectratio=169, 11pt, invertlogo]{ismll-slides}
% - logo=[all, frametitle, headline, none]
% - logotitlepage=[true, false]
% - invertlogo: where to use white instead of black text (on transparent background)
% - nopdfpagenumbers: disable adding a number for each frame to pdf toc

%%%%%%%%%%%%%%%%
% BEAMER THEME %
%%%%%%%%%%%%%%%%
\usetheme[block=fill, background=light, progressbar=foot, numbering=counter]{metropolis}
%\usetheme{Madrid}
\setbeamertemplate{blocks}[default]
\setbeamercovered{transparent}

%%%%%%%%%%%%%%%%
% BIBLIOGRAPHY %
%%%%%%%%%%%%%%%%
% NOTE: Bibliography should be compiled with biber!
%\bibliography{slides-bibliography}
%\addbibresource{slides-bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGE INFORMATION %
%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Active Learning Benchmark}
\subtitle{Towards Comparable Active Learning}
\date{February 31, 2021}
\institute{Information Systems and Machine Learning Lab (ISMLL)\\Institute for Computer Science \\ University of Hildesheim}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Motivation}
	The DAL result landscape is terrible \\ [2mm]
	Every paper uses different datasets and use-cases \\ [1mm]
	Every paper has a different classification model and training regime \\ [1mm]
	No one-fits-all evaluation metric 
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Competitor Papers}
	Some Benchmark papers for DAL exist already:
	\begin{itemize}
		\item Focus on Image Classification
		\item Focus on best possible classifier performance
		\begin{itemize}
			\item Data Augmentation
			\item Type of Optimizer
			\item Semi-Supervised Learning
		\end{itemize}
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Main Idea}
	We focus on the AL algorithms themselves:
	\begin{itemize}
		\item When we find the "best" algorithm, classification performance will come naturally
		\item AL is lacking reproducible research, not strong classification models (random/uncertainty sampling does a good job already)
		\item We incorporate different domains instead of focusing on images
		\item We reduce the amount of true hyperparameters by design
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Theoretical Problems}
	Some conflicting results from other papers:
	\begin{itemize}
		\item Data augmentations (seemingly) can replace diversity components from AL algorithms (under investigation currently)
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Scope}
	The framework is planned to include
	\begin{itemize}
		\item Three different domains (Tabular, Image, Text)
		\item SOTA baseline methods 
		\item Preprocessed datasets (train/test split and seed set)
		\item Tuned classifiers for each dataset
		\item Set logging and evaluation procedures
		\item (Ensemble Approaches)
		\item (Single Domain and Domain-Transfer use cases)
	\end{itemize}
	The framework will \textbf{not} include
	\begin{itemize}
		\item Batch Active Learning
		\item Elaborate Training (Data Augmentation, etc.)
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Design Decisions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{The AL Cycle}
	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{pics/al_cycle.png}
	\end{figure}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Datasets}
	Each dataset should be preprocessed - The features are normalized, the target are one-hot encoded and \textbf{the seed set should be fixed}

	Datasets should be selected by their "potential"

	Each dataset needs an oracle curve that sets an upper bound performance
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Dataset Potential}
	\begin{figure}
		\centering
		\includegraphics[width=0.7\linewidth]{pics/splice_oracle.png}
	\end{figure}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Classifier}
	The classification model should be governed by the dataset

	There is no need to have the same model for every dataset, as long as the model in question is suited well for the data

	Simpler models with less dynamic behaviour are better - SOTA performance is not so important

	The model needs have tuned hyperparameters for each dataset (full dataset or subset?)

	The expectation is that good AL algorithms will also work well on SOTA models
	%Adam vs. SGD?
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Evaluation Metrics}
	Multiple options are available
	\begin{itemize}
		\item Accuracy* / F1-Score
		\item AUC
		\item Advantage over random sampling*
		\item Regret from the oracle curve
		\item Relative performance (0\% = random / 100\% = oracle)
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Cross Validation}
AL can cross validate in two ways
\begin{itemize}
	\item Changing the seed set
	\item Changing the presented sub-samples of unlabeled points
	\item Changing model initialization
\end{itemize}
Currently, I keep the seed set fixed \\ [2mm]
Is there value in fixing everything? (Model checkpoints, pre-sampling data*, etc.)
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Cross Validation}
	\begin{columns}
		\begin{column}{0.3\linewidth}
			Splice (Tabular): \\
			Random: $0.874 \pm 0.01$ \\
			Margin: $0.885 \pm 0.015$ \\
			Difference: $0.011$
			%Changing the data vs. changing the model initialization
		\end{column}
		\hspace{3mm}
		\begin{column}{0.7\linewidth}
			\includegraphics[width=0.9\linewidth]{pics/splice_performance.png}
		\end{column}
	\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Cross Validation}
	\begin{columns}
		\begin{column}{0.5\linewidth}
			Random Sampling
			\includegraphics[width=\linewidth]{pics/splice_random_convergence.jpg}
		\end{column}
		\hspace{3mm}
		\begin{column}{0.5\linewidth}
			Uncertainty Sampling
			\includegraphics[width=\linewidth]{pics/splice_margin_convergence.jpg}
		\end{column}
	\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Cross Validation}
	\begin{figure}
		\centering
		\includegraphics[width=.7\linewidth]{pics/splice_margin100_convergence.jpg}
		\caption*{Uncertainty Sampling}
	\end{figure}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Hyperparameters: Dataset}
	Budget is set to allow convergence

	Seed set is fixed to 1 point per class unless the cold start problem is to pronounced

	Classifier batch size / LR / regularization is optimized for performance \\
	Image datasets might require a trade-off for speed (Cifar10 currently needs 8 days)
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Hyperparameters: Agent}
	Each AL agent defines its own state space with the information from the environment

	Other hyperparameters for baselines need to be investigated
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{True Hyperparameters}
	This leaves very few "true" hyperparameters: 
	\begin{itemize}
		\item Unlabeled Sample Size
		\item \# of cross validation trials
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Ensemble Approaches}
	Uncertainty Sampling profits a lot from ensembles

	The logical thing is to build the final classifier as an ensemble \\ [7mm]
	%
	How do you compare single model approaches to ensemble approaches?

	Prefix a single evaluation model out of the ensemble?
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Single Domain vs. Transfer}
	Single Domain: Fit AL agent on the same dataset that you evaluate on (not useful in practice)

	Domain Transfer: Fit AL agent on domain with large resources and evaluate agent on domain with few resources

	Often used in NLP (fit agent on English data and evaluate for Spanish / German)
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{The End}
	Cheers
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\begin{frame}[fragile]{Classifier Training}
	\begin{table}
		\centering
		\begin{tabular}{l || l}
			\textbf{General Setup} & \textbf{Speed Optimizations} \\
			\hline
			Cross-Entropy Loss & Fine tuning \\ [1mm]
			No case weights & Aggressive Early Stopping \\[1mm]
			At least one epoch & \\[1mm]
			NAdam optimizer &
		\end{tabular}
	\end{table}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Results for Splice}
	\begin{table}
		\centering
		\begin{tabular}{l || c | c }
			\textbf{Agent} & \textbf{Acc} & \textbf{AUC} \\
			\hline
			Oracle & 0.8806 $\pm$ 0.0141 & 0.8530 $\pm$ 0.0132 \\
			SAR & 0.8839 $\pm$ 0.0142 & 0.8339 $\pm$ 0.0118 \\
			Margin & 0.8849 $\pm$ 0.0112 & 0.8309 $\pm$ 0.0141 \\
			Entropy & 0.8822 $\pm$ 0.0156 & 0.8285 $\pm$ 0.0154 \\
			Coreset & 0.8788 $\pm$ 0.0105 & 0.8246 $\pm$ 0.0108 \\
			Random & 0.8744 $\pm$ 0.0144 & 0.8152 $\pm$ 0.0129
		\end{tabular}
		\caption*{Results for Splice}
	\end{table}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}