### 2024-10-10 doc/rebuttal.md
camera ready

# Reviewer iCXF
rating: 7, confidence 3
```
Thank you for your review...
```

The results on the synthetic benchmarks are unsurprising and the contribution is unclear since they show that methods that focus on uncertainty sampling like margin fail on distributions that are adversarially designed to make them fail.
```
Our primary goal is to validate these approaches through a targeted ablation study, highlighting the strengths and limitations of both types of algorithms. 
You are correct when you say that algorithms like margin sampling behave exactly as expected, but in our opinion, the behavior of a hybrid algorithm like BADGE was completely untested in these adversarial environments (in the case of BADGE, we revealed weaknesses akin to vanilla uncertainty sampling methods, even though BADGE uses an clustering-like algorithm to select its samples).
These datasets mainly serve as a resource for benchmarking future AL methods, particularly for hybrid approaches that use ideas from uncertainty sampling and clustering/diversity methods.
```

The algorithm for the greedy oracle has no clear guarantees, the design choices are unclear and the description is hard to follow.
```
We designed our oracle based in the following two ideas:
- Have an algorithm that does not rely on search or excessive enumeration, so it can be computed efficiently 
- Pose a performance that is "good enough", so that it beats every AL algorithm on average without wasting computational resources for diminishing lifts

Since we are mainly interested in beating other AL algorithms and not a theoretically optimal upper bound we did not see the need derive a performance guarantee for our oracle.

To improve the description of the oracle, we will add comments to the pseudocode in Appendix H and reference it more clearly in the main body:
Remove Line 258: The pseudocode for our oracle can be found in App. H.
Add: For a detailed description of the oracle algorithm and how it is used in the AL loop, please refer to App. H.
```

The evaluation is done with relatively small models and datasets and it is unclear whether it translates to larger ones.
```
We agree that no comparable benchmark has shown a systematic conversion of small-model setup to large-model setups.
However, we have evidence in domain-specific literature that supports our hypothesis that our ranking transfers to larger models:
[4] use larger image models like ViT-B32 and also find least confidence sampling and BADGE to be the best algorithms for images
[5] use BERT models in the text domain and also find BADGE to be the best (they don't evaluate margin sampling)
[6] use considerably larger MLPs for the tabular domain than we do, but also find margin sampling to be consistenly top-performing

References:
[4]: Zhang, Jifan, et al. "LabelBench: A Comprehensive Framework for Benchmarking Adaptive Label-Efficient Learning." Journal of Data-centric Machine Learning Research (2024).
[5]: Rauch, Lukas, et al. "Activeglae: A benchmark for deep active learning with transformers." Joint European Conference on Machine Learning and Knowledge Discovery in Databases. Cham: Springer Nature Switzerland, 2023.
[6]: Bahri, Dara, et al. "Is margin all you need? An extensive empirical study of active learning on tabular data." arXiv preprint arXiv:2210.03822 (2022).
```

# Reviewer 3W2R
rating: 7, confidence 3
```
Thank you for your review...
```
However this work should probably cite and discuss Zhang et al.
```
We will add Zhang et al to Table 1 and our Related Work
```

The "oracle" needs to be explained earlier. It is discussed and referenced a lot, but you don't actually describe what it is until section 6. I was very confused, since typically "oracle" might also refer to an annotator. Things made sense at section 6, but a preview of what exactly the oracle does should be discussed as soon as possible. Looking again, it is discussed at line 53 but I actually found this paragraph very vague.
```
We will insert a sentence in line 55 to clarify the main idea of our oracle algorithm early on:
"Our oracle relies on directly testing a small sample of points in every iteration if they induce an improvement in test accuracy and selects the optimal point from that small sample."
```

what is "sngl" in Table 1? Single point sampling per round?
```
Yes, we will improve the clarity in Table 1
```

what is 9(14) in Table 1 in the last row?
```
We have 9 datasets in our benchmark, 5 of which have a pre-encoded version (excluding text and synthetic) , which brings the total number of experiments to 14.
We will include this description in the caption of Table 1.
```

not sure what i\in means really after line 93
```
It simply indexes all points in the unlabeled set.
We will improve the clarity here as well
```

I think Figure 1 needs better descriptions and a better legend. I think I get the main point but it's confusing. E.g., "True mean" is vague, what is that the true mean of? Purple curve doesn't even appear on the legend.
```
We will add all visible curves to the legend in Figure 1.
Additionally, we will revise the caption of Figure 1 to make it more accessible for the readers.
```

the choice of validation on the entire dataset needs more discussion (line 194). This is a huge criticism of active learning research, to choose parameters based on a full validation set. I understand the argument for lower variance in research evaluation, but I don't think the justification here is sufficient. In particular, Figure 1 argues that the high variance in research results is a problem. Why mask it, with an unrealistic validation? Won't that make things worse?
```
It is true that, by proposing to use a fully labeled validation set, we implicitly claim that most AL research so far had been flawed.
However, we don't want to challenge AL literature as a whole, but rather push for a stronger separation of AL research (with validation set) and AL applications (without validation set).
With our approach, we are not trying to mask the high variance; instead, we argue that **due** to the high variance, we should fully optimize our hyperparameters on validation to avoid exacerbating its effects.
The core hypothesis is that a top-performing algorithm in an experiment with good hyperparameters also performs well in the application case with bad/worse hyperparameters.
```

# Reviewer x7yL
rating: 4, confidence 5

```
Thank you for your review...
```
Although the author states that the proposed benchmark is the first one that contains multiple domains, however, in each domain, many related empirical studies focus on specific domain, e.g., [r1], [r2], [r3], which weakens the contribution of this paper.
```
We will include [r1] into Table 1 and our related work ([r2] and [r3] do not provide experiments on their own).
AL benchmarks are still riddled with comparability issues, because their choice of datasets, domains and models often is completely disjoint ([r2] Sec. 4.1 and [r3] Tab. 1).
In our opinion, we need a benchmark that spans all domains and the most commonly used datasets and models to fix this problem.
As described in our Section 5, we selected our datasets for maximum overlap with previous works. The same is true for our model selection, which either uses common models like ResNet18 or trivial to reproduce models, like MLPs.
[r2]: Zhang, Zhisong, Emma Strubell, and Eduard Hovy. "A survey of active learning for natural language processing." arXiv preprint arXiv:2210.10109 (2022).
[r3] Schröder, Christopher, and Andreas Niekler. "A survey of active learning for text classification using deep neural networks." arXiv preprint arXiv:2008.07267 (2020).
```

The insights (e.g., the importance of multiple runs) build on existing knowledge in the active learning field.
```
Even though it is common knowledge that repeating an experiment enough times is crucial for ML research, in the field of Active Learning, we still observe many authors not doing so. 
To the best of our knowledge, we are the first to provide insight into exactly how often an AL experiment should be repeated in order to generate reliable results.
The same is true for experimenting across different domains. Everyone knows that algorithms should be tested in as many domains as possible, but few actually apply that knowledge.
```

This paper takes too much space on the greedy oracle algorithm and seeding strategies, which may be complex for some readers to fully get the meaning.
```
See below
```

In the main manuscript, the author takes too much space to discuss why they need to have such experimental settings and how they set the experiments. The discussions of the experimental results are weak. Maybe this paper is more suitable to submit to a journal.
```
As you have mentioned in your review, there are a fair number of benchmarks for AL already.
Each of those focus on 1-2 domains and place high emphasis on their results.
However, apart from [1], each benchmark has the experimental setup and technical details on low priority, leading to a situation where everyone is doing their own framework in isolation and an informed discussion about best practices is very difficult.
We hope (in combination with [1]) to open the door for such discussion in the area of AL.
Nonetheless, we will extend discussion of our results. 
Based on the comment from another reviewer, we plan to study our results from the dimension of class imbalance, which currently is missing in our work. 
[1]: Ji, Yilin, et al. "Randomness is the root of all evil: more reliable evaluation of deep active learning." Proceedings of the IEEE/CVF Winter Conference on Applications of Computer Vision. 2023.
```

"In semi-supervised learning, which is to train a fixed encoder-model, pre-encode the datasets and then only train a single linear layer as classifier." It can be computationally efficient, but it will not fully exploit the strengths of deep learning in an active learning scenario. A more typical and potentially more effective approach in deep active learning involves training the feature extractor alongside the classifier to ensure that the representations are continuously adapted to the newly labeled data.
```
As described in Section 5.2, we are not interested in beating the SOTA performance on any dataset, but rather in providing a fast and low-variance framework.
The same idea applies to our choice of semi-supervised learning.
For that reason, we opted for pre-encoded datasets instead of more sophisticated approaches.
Finally, the authors of [4] provide good evidence that the ranking of algorithms does not change when the semi-supervised training changes (at least for the image domain).
[4]: Zhang, Jifan, et al. "LabelBench: A Comprehensive Framework for Benchmarking Adaptive Label-Efficient Learning." Journal of Data-centric Machine Learning Research (2024).
```

# Reviewer c7pZ
rating: 5, confidence 5
```
Thank you for your review...
```

Include more related benchmarks. For example [1] studies extensively the behavior of active learning in tabular datasets. LabelBench [2] has also proposed using embedded features and semi-supervised learning as part of their benchmark. While the authors in this paper conjecture "a well-performing method in our benchmark will also generalize well to larger datasets and classifiers", LabelBench has already demonstrated this to be true. In addition, margin sampling performing well on tabular datasets is also found in [1].
```
[1] and [2] will be integrated into Table 1 and our related work section.
Based on the comment of another reviewer we aim to extend our discussion of our results in the appendix, which will also include a more in-depth comparison of our results with other benchmarks like [1].
```

The proposed metric of average ranking may not be a convenient/intuitive metric for benchmarks. Specifically, whenever a new algorithm is introduced, the scores of every algorithm will change. Moreover, practitioners are generally interested in either accuracy or label-efficiency (number of labels needed to reach a certain accuracy). The adopted metric does not capture any of these quantities directly. In some cases, entropy, badge and margin may all perform very similarly in accuracy/label-efficiency but remains stable in their relative ranking. The metric in this paper would significantly penalize the algorithms that perform relatively worse, which does not give an accurate quantification in the actual performance of the algorithms themselves.
```
We fully agree with the reviewer that accuracy scores need to be reported for any classification problem.
We have all accuracy scores in Appendix K, including standard deviations and the amount of wins per algorithm.
We would like to point out that our benchmark aggregates results from different datasets, which we cannot be done by simply averaging the accuracies to display them in a table, etc.
For a fair comparison, we rely on the paired-t test, which the Critical-Difference-Diagrams conveniently provide.
Lastly, we were unable to include a second results table besides Table 3 within the page limit and therefore located our accuracy values in the Appendix.
```

An interesting finding in LabelBench is imbalanced active learning algorithms like GALAXY performs significantly better on imbalanced datasets. Since most of the language and tabular datasets are imbalanced, it would be interesting to see how such algorithms perform. The conclusion that "entropy sampling" or "margin sampling" is the best may not be entirely accurate.
```
We omitted information about class imbalance in our datasets, as we did not focus on that aspect.
However, both our text and 2 out of 3 tabular datasets are imbalanced by nature. Implementing GALAXY and discussing results along this dimension as well is a great suggestion. 
Since testing GALAXY on our entire benchmark is a significant computational effort, we can only provide experiments on the TopV2 dataset for this rebuttal (doc/img/micro_dna).
However, GALAXY is indeed the top-performing algorithm on the imbalanced TopV2 dataset.
We thank the reviewer for this valuable addition to the benchmark and will fully incorporate GALAXY for the camera-ready version.
```